---
title: "Predict depression from FH indicators"
output: html_document
date: "2024-27-03"
---

This script contains
1) full assessment of the performance of the indicators in predicting depression
2) modified list of assessment results for Figure 4, including combined models
Input: Scripts/setup.R, indicators_PRS_index.RData
Output: assessment.RData, assessment_display4.RData, assessment_display4_full_results.csv
Status: Running

```{r setup, include=FALSE}	
# Clear workspace
rm(list = ls())
getwd()
setwd("LT-FH")
source(file = "Scripts/setup.R")

# Load data
load(file = here("Int_results", "indicators_PRS_index.RData"))
head(indicators_PRS_index)
# Narrow to sample with PRS, turn DEP_index into factor
indicators_PRS_index <- indicators_PRS_index |>
  pivot_wider(names_from = indicator, values_from = value) |>
  filter(!is.na(PRS)) |>
  mutate(DEP_index = as.factor(DEP_index)) |>
  pivot_longer(
    cols = c("plusminus", "proportion", "FLS", "PAFGRS", "PAFGRSplus", "PRS"),
    names_to = "indicator",
    values_to = "value"
  )


# Parameters
index_disorder <- "DEP"
index_labels <- paste("index", index_disorder, sep = "_")
```


# Prediction of depression
```{r}
# Run with list indicators_PRS_index instead
# Initiate empty list to store results
assessment <- list()
# Parameters
K <- Ks[index_disorder]

for (type in all_types) {
  # Select dataframe from list
  # indicators_df <- indicators_PRS_index[[type]]
  indicators_df <- indicators_PRS_index |>
    filter(coding == type) |>
    pivot_wider(names_from = indicator, values_from = value)

  print(head(indicators_df))
  # Assess performance for all 5 indicators in predicting depression
  assessment[[type]] <- sapply(predictors, FUN = function(x) {
    # Set input arguments
    diagnosis <- indicators_df[["DEP_index"]]
    str(diagnosis)
    # Set predictor
    predictor <- indicators_df[[x]]
    str(predictor)

    # Run assessment function
    assess_performance_dichot(
      true_status = as.numeric(diagnosis), predictor = predictor, K
    )
  })
}

# Check
assessment
```

# Assessment for main results (display item 4)

```{r}
# AUC of combined model
# Combine PAFGRSplus and PRS

# Example usage:
indicators_df <- indicators_PRS_index |>
  mutate(indicator = as.factor(indicator))

head(indicators_df)
table(indicators_df$DEP_index, useNA = "ifany")
table(indicators_df$indicator, useNA = "ifany")

# Assess performance for different predictors
results <- list(
  assess_performance_flexibly(indicators_df, coding_type = "DEP_simple", outcome = "DEP_index", predictors = c("plusminus")),
  assess_performance_flexibly(indicators_df, coding_type = "DEP_simple", outcome = "DEP_index", predictors = c("proportion")),
  assess_performance_flexibly(indicators_df, coding_type = "DEP_simple", outcome = "DEP_index", predictors = c("FLS")),
  assess_performance_flexibly(indicators_df, coding_type = "DEP_simple", outcome = "DEP_index", predictors = c("PAFGRS")),
  assess_performance_flexibly(indicators_df, coding_type = "DEP_simple", outcome = "DEP_index", predictors = c("PAFGRSplus")),
  assess_performance_flexibly(indicators_df, coding_type = "DEP_simple_with_ANX_simple", outcome = "DEP_index", predictors = c("PAFGRSplus")),
  assess_performance_flexibly(indicators_df, coding_type = "DEP_simple_with_ANX_simple", outcome = "DEP_index", predictors = c("PRS")),
  assess_performance_flexibly(indicators_df, coding_type = "DEP_simple_with_ANX_simple", outcome = "DEP_index", predictors = c("PAFGRSplus", "PRS"))
)

# Convert results list to dataframe
results_list <- map(results, "results")
results_list
results_df <- do.call(rbind, lapply(results_list, as.data.frame))
rownames(results_df) <- NULL
results_df

# Double check model coefficients
model_list <- map(results, "model")
lapply(model_list, summary) # yes all fine, correspond to results_df

# Compare roc objects of PAFGRS+Sex/Age and PAFGRS+Sex/Age/ANX
roc_obj_list <- map(results, "roc_obj")
roc1 <- roc_obj_list[[5]] # PAFGRS+Sex/Age
roc2 <- roc_obj_list[[6]] # PAFGRS+Sex/Age/ANX
roc.test(roc1, roc2) # Show's that the difference likely is different from 0
```

# Repeat assessment with scaled indicators for interpretable coefficients
```{r}
indicators_df <- indicators_PRS_index |>
  mutate(indicator = as.factor(indicator))
head(indicators_df)

indicators_scaled <- indicators_df |>
  pivot_wider(names_from = c(coding, indicator), values_from = value) |>
  mutate(across(.cols = c(ends_with("plusminus"), ends_with("proportion"), ends_with("FLS"), ends_with("PAFGRS"), ends_with("PAFGRSplus")),
                .fns = ~ as.numeric(scale(.)))) |>
  pivot_longer(cols = -c(pident, DEP_index), names_to = c("coding", "indicator"), values_to = "value",
               names_sep = "_(?=[^_]+$)")  # split at the final underscore
head(indicators_scaled)
write.csv(indicators_scaled, file = here("Int_results", "indicators_scaled.csv"), row.names = FALSE)

temp <- indicators_scaled |>
  filter(coding == "DEP_simple") |>
  pivot_wider(names_from = indicator, values_from = value)
head(temp)

summary(as.numeric(temp$plusminus))
summary(as.numeric(temp$proportion))
summary(as.numeric(temp$FLS))
summary(as.numeric(temp$PAFGRS))
summary(as.numeric(temp$PAFGRSplus))
summary(as.numeric(temp$PRS))

sd(as.numeric(temp$plusminus))
sd(as.numeric(temp$proportion))
sd(as.numeric(temp$FLS))
sd(as.numeric(temp$PAFGRS))
sd(as.numeric(temp$PAFGRSplus))
sd(as.numeric(temp$PRS))

plot(density(as.numeric(temp$plusminus), na.rm = TRUE), main = "Density plot of Scaled PlusMinus")
plot(density(as.numeric(temp$proportion), na.rm = TRUE), main = "Density plot of Scaled Proportion")
plot(density(as.numeric(temp$FLS), na.rm = TRUE), main = "Density plot of Scaled FLS")
plot(density(as.numeric(temp$PAFGRS), na.rm = TRUE), main = "Density plot of Scaled PAFGRS")
plot(density(as.numeric(temp$PAFGRSplus), na.rm = TRUE), main = "Density plot of Scaled PAFGRSplus")
plot(density(rnorm(1e4)))

# Run assessment
# Assess performance for different predictors
results_scaled <- list(
  assess_performance_flexibly(indicators_scaled, coding_type = "DEP_simple", outcome = "DEP_index", predictors = c("plusminus")),
  assess_performance_flexibly(indicators_scaled, coding_type = "DEP_simple", outcome = "DEP_index", predictors = c("proportion")),
  assess_performance_flexibly(indicators_scaled, coding_type = "DEP_simple", outcome = "DEP_index", predictors = c("FLS")),
  assess_performance_flexibly(indicators_scaled, coding_type = "DEP_simple", outcome = "DEP_index", predictors = c("PAFGRS")),
  assess_performance_flexibly(indicators_scaled, coding_type = "DEP_simple", outcome = "DEP_index", predictors = c("PAFGRSplus")),
  assess_performance_flexibly(indicators_scaled, coding_type = "DEP_simple_with_ANX_simple", outcome = "DEP_index", predictors = c("PAFGRSplus")),
  assess_performance_flexibly(indicators_scaled, coding_type = "DEP_simple_with_ANX_simple", outcome = "DEP_index", predictors = c("PRS")),
  assess_performance_flexibly(indicators_scaled, coding_type = "DEP_simple_with_ANX_simple", outcome = "DEP_index", predictors = c("PAFGRSplus", "PRS"))
)
# Convert results list to dataframe
results_list_scaled <- map(results_scaled, "results")
results_list_scaled
results_df_scaled <- do.call(rbind, lapply(results_list_scaled, as.data.frame))
rownames(results_df_scaled) <- NULL
results_df_scaled

# Add column with ORs per SD increase
results_df_scaled <- results_df_scaled |>
  mutate(
    OR = exp(coeff),
    ci_lower = exp(coeff - 1.96 * coeff.SE), # Lower bound of CI
    ci_upper = exp(coeff + 1.96 * coeff.SE) # Upper bound of CI
  ) 
results_df_scaled
  
results_df_scaled_for_table <- results_df_scaled |>
  mutate(AUC = round(AUC, 3), `SE (AUC)` = round(SE, 3), Coefficient = round(coeff, 3), `SE (Coefficient)` = round(coeff.SE, 3), 
  OR = round(OR, 2), `95% CI Lower` = round(ci_lower, 2), `95% CI Upper` = round(ci_upper, 2), `P-value` = pvalue, .keep = "unused") 
```


# Save

```{r}
# Save results for main display item 4
assessment_display4 <- results_df
save(assessment_display4, file = here("Int_results", "assessment_display4.RData"))

# Save results_df for full results table in supplement
assessment_display4_full_results <- results_df
write.csv(
  assessment_display4_full_results,
  file = here("Results", "assessment_display4_full_results.csv"),
)

# Save complete results as list
save(assessment, file = here("Results", "assessment.RData"))
assessment

# Save each dataframe from complete results as .csv
for (type in types) {
  write.csv(
    assessment[[type]],
    file = here("Results", paste0("assessment_", type, ".csv")),
    row.names = TRUE
  )
}

# Save results_df
save(results_df_scaled, file = here("Int_results", "results_df_scaled.RData"))
write.csv(results_df_scaled_for_table, file = here("Int_results", "results_scaled_table.csv"))
```