---
title: "FH indicators DEP/ANX"
output: html_document
date: "2024-07-02"
---

This script
1) prepares the dataframes for family history indicator calculation by separating for assessment type  
2) calculates the family history indicators for depression only, and depression + anxiety  
3) prepares a file with pident, PRS and FH indicators, excluding related individuals
Input: Scripts/indicator_functions.R, joint_fam_work_R2.csv, DAP2406_PRS_pident.sav, NESDA_lifetime_depression_cleaned.csv, FTI_famsize.csv, FTI_age.csv, FTI_diagnoses.csv
Output: Int_results/FH.RData, Int_results/indicators_PRS_index.RData, Int_results/indicators_with_index.RData
Status: Running
!!! Contains pidents of related individuals


- preparation of dataframes for indicator calculation
- calculation of the FH indicators for DEP, and DEP+ANX
- preparation of file with pident, PRS and FH indicators
- generation of descriptives of the final sample (FTI and PRS available)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

setwd("LT-FH")
getwd()
# Clear workspace
rm(list = ls())

# Setup - packages and functions
source(file = "Scripts/indicator_functions.R")

# NESDA which participants are related
NESDA_related <- read.csv(here("Data", "joint_fam_work_R2.csv"))
# NESDA PRS with pident added by data management
PRS <- read_sav(file = here("Data", "DAP2406_PRS_pident.sav"))
# NESDA lifetime diagnoses (with controls cleaned for bipolar and anxiety)
index_diagnoses <- read.csv(here("Data", "NESDA_lifetime_depression_cleaned.csv"))
# read fti data
FTI_famsize <- read.csv(here("Data", "FTI_famsize.csv"))
FTI_age <- read.csv(here("Data", "FTI_age.csv"))
FTI_diagnoses <- read.csv(here("Data", "FTI_diagnoses.csv"))

# How many relatives were missing age information? 182/8005 = 2.27%
dim(FTI_age)
sum(is.na(FTI_age$age))
182/8005
```

# Data prep

```{r }
# Get relevant columns from FTI data
FH_full <- FTI_diagnoses |>
    dplyr::select(pident, fammember, starts_with("depdiag"), starts_with("anxdiag")) |>
    pivot_wider(
        id_cols = c(pident),
        names_from = fammember,
        values_from = c(depdiag_simple:anxdiag_2020paper)
    )

# Merge index diagnoses with FTI diagnoses
# remove those with missing DEP_index # I don't remember why this was done, becomes more complicated when including ANX
index_diagnoses <- index_diagnoses |>
    rename("DEP_index" = "caseDEP") |>
    rename("ANX_index" = "ever_anx")
head(index_diagnoses)

FH <- index_diagnoses[c("pident", "DEP_index", "ANX_index")] |>
    inner_join(FH_full, by = "pident")
dim(FH)

# Final sample pidents to align all parameters
current_sample <- FH$pident

# Prepare sex and age data
# Sex is derived from recorded type of family member (mother vs. father, brother vs. sister)
total_rows <- length(current_sample)
sex_df <- as.data.frame(cbind(
    mo = rep("F", total_rows),
    fa = rep("M", total_rows),
    br1 = rep("M", total_rows),
    si1 = rep("F", total_rows),
    br2 = rep("M", total_rows),
    si2 = rep("F", total_rows),
    br3 = rep("M", total_rows),
    si3 = rep("F", total_rows),
    br4 = rep("M", total_rows),
    si4 = rep("F", total_rows),
    si5 = rep("F", total_rows),
    si6 = rep("F", total_rows),
    br5 = rep("M", total_rows),
    br6 = rep("M", total_rows),
    br7 = rep("M", total_rows),
    br8 = rep("M", total_rows),
    br9 = rep("M", total_rows),
    si7 = rep("F", total_rows),
    si8 = rep("F", total_rows),
    si9 = rep("F", total_rows),
    si10 = rep("F", total_rows)
))

# Age was calculated in FTI_preprocessing, here just adapt dataframe
age_df <- FTI_age |>
    pivot_wider(
        id_cols = c(pident),
        names_from = c(fammember),
        values_from = c(age)
    ) |>
    filter(pident %in% current_sample) |> # make sure only the correct pidents are used. order should remain the same.
    dplyr::select(-pident)
dim(age_df)
# Changing 0s to 1s because PAFGRSplus cannot handle zeros
age_df[!is.na(age_df) & age_df == 0] # 43 zeros.
age_df[!is.na(age_df) & age_df == 1] # 8 ones
# Confusion fix: Replace zeros with NA
age_df[age_df == 0] <- 1 # replace zeros with ones, resulting in 51 ones
age_df[!is.na(age_df) & age_df == 0]

famsize <- FTI_famsize |>
    mutate(famsize = sib_nr + 2) |>
    filter(pident %in% current_sample) |> # make sure only the correct pidents are used. order should remain the same.
    dplyr::select(famsize)
famsize <- as.vector(famsize$famsize)
length(famsize)
```



# Create list of dataframes (one dataframe per by type of diagnostic data)
```{r}
# Here we fill the list FTI_data in an orderly manner. The process is messy but the resulting list works.
# Initiate empty list to store separated dataframes
FTI_data <- list()

# Family diagnosis data, elaborate coding
FTI_data[["DEP_elaborate"]] <- FH |>
    dplyr::select(starts_with("depdiag_elaborate") & !contains("0imputed")) |> # not yet flipped! choice of diagnostic coding
    rename_with(~ gsub("depdiag_elaborate", "DEP", .x), .cols = everything()) |>
    # remove columns of br8 and br9 as they are all NA
    mutate(DEP_br8 = NULL, DEP_br9 = NULL)
# Rename columns
colnames(FTI_data[["DEP_elaborate"]]) <- sapply(colnames(FTI_data[["DEP_elaborate"]]), flip_colnames) # flip column names around _ to match FH functions

# Family diagnosis data, simple coding
FTI_data[["DEP_simple"]] <- FH |>
    dplyr::select(starts_with("depdiag_simple") & !contains("0imputed")) |> # not yet flipped! choice of diagnostic coding
    rename_with(~ gsub("depdiag_simple", "DEP", .x), .cols = everything()) |>
    # remove columns of br8 and br9 as they are all NA
    mutate(DEP_br8 = NULL, DEP_br9 = NULL)
# Rename columns
colnames(FTI_data[["DEP_simple"]]) <- sapply(colnames(FTI_data[["DEP_simple"]]), flip_colnames) # flip column names around _ to match FH functions

# Family diagnosis data, elaborate coding, zero imputed
FTI_data[["DEP_elaborate_0imputed"]] <- FH |>
    dplyr::select(starts_with("depdiag_elaborate_0imputed")) |> # not yet flipped! choice of diagnostic coding
    rename_with(~ gsub("depdiag_elaborate_0imputed", "DEP", .x), .cols = everything()) |>
    # remove columns of br8 and br9 as they are all NA
    mutate(DEP_br8 = NULL, DEP_br9 = NULL)
# Rename columns
colnames(FTI_data[["DEP_elaborate_0imputed"]]) <- sapply(colnames(FTI_data[["DEP_elaborate_0imputed"]]), flip_colnames) # flip column names around _ to match FH functions

# Family diagnosis data, simple coding, zero imputed
FTI_data[["DEP_simple_0imputed"]] <- FH |>
    dplyr::select(starts_with("depdiag_simple_0imputed")) |> # not yet flipped! choice of diagnostic coding
    rename_with(~ gsub("depdiag_simple_0imputed", "DEP", .x), .cols = everything()) |>
    # remove columns of br8 and br9 as they are all NA
    mutate(DEP_br8 = NULL, DEP_br9 = NULL)
# Rename columns
colnames(FTI_data[["DEP_simple_0imputed"]]) <- sapply(colnames(FTI_data[["DEP_simple_0imputed"]]), flip_colnames) # flip column names around _ to match FH functions

# Family diagnosis data, 2020 paper coding
FTI_data[["DEP_2020paper"]] <- FH |>
    dplyr::select(starts_with("depdiag_2020paper")) |> # not yet flipped! choice of diagnostic coding
    rename_with(~ gsub("depdiag_2020paper", "DEP", .x), .cols = everything()) |>
    # remove columns of br8 and br9 as they are all NA
    mutate(DEP_br8 = NULL, DEP_br9 = NULL)
# Rename columns
colnames(FTI_data[["DEP_2020paper"]]) <- sapply(colnames(FTI_data[["DEP_2020paper"]]), flip_colnames) # flip column names around _ to match FH functions


# Family diagnosis data, elaborate coding
FTI_data[["DEP_elaborate_with_ANX_elaborate"]] <- FH |>
    dplyr::select(
        starts_with("depdiag_elaborate") & !contains("0imputed"),
        starts_with("anxdiag_elaborate") & !contains("0imputed")
    ) |> # not yet flipped! choice of diagnostic coding
    rename_with(~ gsub("depdiag_elaborate", "DEP", .x), .cols = everything()) |>
    rename_with(~ gsub("anxdiag_elaborate", "ANX", .x), .cols = everything()) |>
    # remove columns of br8 and br9 as they are all NA
    mutate(DEP_br8 = NULL, DEP_br9 = NULL, ANX_br8 = NULL, ANX_br9 = NULL)
# Rename columns
colnames(FTI_data[["DEP_elaborate_with_ANX_elaborate"]]) <- sapply(colnames(FTI_data[["DEP_elaborate_with_ANX_elaborate"]]), flip_colnames) # flip column names around _ to match FH functions

# Family diagnosis data, simple coding
FTI_data[["DEP_simple_with_ANX_simple"]] <- FH |>
    dplyr::select(
        starts_with("depdiag_simple") & !contains("0imputed"),
        starts_with("anxdiag_simple") & !contains("0imputed")
    ) |> # not yet flipped! choice of diagnostic coding
    rename_with(~ gsub("depdiag_simple", "DEP", .x), .cols = everything()) |>
    rename_with(~ gsub("anxdiag_simple", "ANX", .x), .cols = everything()) |>
    # remove columns of br8 and br9 as they are all NA
    mutate(DEP_br8 = NULL, DEP_br9 = NULL, ANX_br8 = NULL, ANX_br9 = NULL)

# Rename columns
colnames(FTI_data[["DEP_simple_with_ANX_simple"]]) <- sapply(colnames(FTI_data[["DEP_simple_with_ANX_simple"]]), flip_colnames) # flip column names around _ to match FH functions

# Check
lapply(FTI_data, head)
```


# Calculate FH indicators for index disorder depression
```{r }
indicators_dep <- list()
indicators_with_index_dep <- list()
# One loop for indicators based on family history of depression only
for (type in types) {
    indicators_dep[[type]] <- score_FH(FTI_data[[type]],
        methods = c("plusminus", "proportion", "PAFGRS", "PAFGRSplus", "FLS"),
        family_sizes = famsize, sexes = sex_df, ages = age_df,
        disorders = c("DEP"), index_disorder = "DEP", disorder_FLS = "DEP",
        Ks = Ks["DEP"], h2ls = h2ls["DEP"], rgs = rg_DEP.ANX,
        threshold_list = disorder_thresholds, K_relative = K_relative, K_sporadic = K_sporadic,
        AO_lower = AO_lower, AO_upper = AO_upper
    )
    # for certain plots, add index individual's diagnostic status
    indicators_with_index_dep[[type]] <- cbind(FH["DEP_index"], indicators_dep[[type]])
}

# One loop for indicators based on family history of depression and anxiety, items added to same list as above
for (type in types_ANX) {
    indicators_dep[[type]] <- score_FH(FTI_data[[type]],
        methods = c("plusminus", "proportion", "PAFGRS", "PAFGRSplus", "FLS"),
        family_sizes = famsize, sexes = sex_df, ages = age_df,
        disorders = c("DEP", "ANX"), index_disorder = "DEP", disorder_FLS = "DEP.ANX",
        Ks = Ks[c("DEP", "ANX")], h2ls = h2ls[c("DEP", "ANX")], rgs = rg_DEP.ANX,
        threshold_list = disorder_thresholds, K_relative = K_relative, K_sporadic = K_sporadic,
        AO_lower = AO_lower, AO_upper = AO_upper
    )
    # for certain plots, add index individual's diagnostic status
    indicators_with_index_dep[[type]] <- cbind(FH["DEP_index"], indicators_dep[[type]])
}

head(indicators_with_index_dep[["DEP_elaborate"]])
str(indicators_with_index_dep)

```

# Calculate FH indicators for index disorder anxiety

```{r }
indicators_anx <- list()
indicators_with_index_anx <- list()

# Prep data for ANX Simple indicators (NO dep)
FTI_anx <- FTI_data[["DEP_simple_with_ANX_simple"]] |>
    dplyr::select(ends_with("ANX"))
# One loop for indicators based on family history of anxiety only
for (type in "ANX_simple") {
    indicators_anx[[type]] <- score_FH(FTI_anx,
        methods = c("plusminus", "proportion", "PAFGRS", "PAFGRSplus", "FLS"),
        family_sizes = famsize, sexes = sex_df, ages = age_df,
        disorders = c("ANX"), index_disorder = "ANX", disorder_FLS = "ANX",
        Ks = Ks["ANX"], h2ls = h2ls["ANX"], rgs = rg_DEP.ANX,
        threshold_list = disorder_thresholds, K_relative = K_relative, K_sporadic = K_sporadic,
        AO_lower = AO_lower, AO_upper = AO_upper
    )
    # for certain plots, add index individual's diagnostic status
    indicators_with_index_anx[[type]] <- cbind(FH["ANX_index"], indicators_anx[[type]])
}
str(indicators_anx)

# One loop for indicators based on family history of depression and anxiety, items added to same list as above
# draws input from FTI_data
for (type in "DEP_simple_with_ANX_simple") {
    indicators_anx[[type]] <- score_FH(FTI_data[[type]],
        methods = c("plusminus", "proportion", "PAFGRS", "PAFGRSplus", "FLS"),
        family_sizes = famsize, sexes = sex_df, ages = age_df,
        disorders = c("DEP", "ANX"), index_disorder = "ANX", disorder_FLS = "DEP.ANX",
        Ks = Ks[c("DEP", "ANX")], h2ls = h2ls[c("DEP", "ANX")], rgs = rg_DEP.ANX,
        threshold_list = disorder_thresholds, K_relative = K_relative, K_sporadic = K_sporadic,
        AO_lower = AO_lower, AO_upper = AO_upper
    )
    # for certain plots, add index individual's diagnostic status
    indicators_with_index_anx[[type]] <- cbind(FH["ANX_index"], indicators_anx[[type]])
}
```

# Checks
```{r}
add_pident <- indicators_with_index_dep[["DEP_elaborate"]] |>
        mutate(pident = FH$pident) |>
        mutate(coding = type)



head(indicators_anx[["DEP_simple_with_ANX_simple"]])
head(indicators_dep[["DEP_simple_with_ANX_simple"]])

head(indicators_with_index_anx[["DEP_elaborate_with_ANX_elaborate"]])
head(indicators_with_index_dep[["DEP_simple_with_ANX_simple"]])
str(indicators_with_index_anx)
```


# Prep file with pident and PRS (index disorder depression)
Exclusion steps:
- Exclude relatives with PRS missing
- Exclude one index individual per related pair (known siblings in NESDA)

```{r}
# Format PRS df
# This is only data for genotyped individuals, so naturally there are no NAs
PRS <- PRS |>
    dplyr::select(pident, SCORE1_SUM) |>
    mutate(PRS = c(scale(SCORE1_SUM)), .keep = "unused")
head(PRS)
# Add pident to indicators_with_index and merge with PRS
indicators_PRS_index <- list()
for (type in all_types) {
    add_pident <- indicators_with_index_dep[[type]] |>
        mutate(pident = FH$pident) |>
        mutate(coding = type)
    # Merge with PRS. The default is an inner join, where only matching pidents are maintained.
    indicators_PRS_index[[type]] <- merge(
        x = add_pident,
        y = PRS,
        by = "pident"
    )
}

add_pident <- indicators_with_index_dep[["DEP_elaborate"]] |>
        mutate(pident = FH$pident) |>
        mutate(coding = type)
head(add_pident)

sapply(indicators_PRS_index, function(x) sum(is.na(x$PRS))) # there should be no NAs
sapply(indicators_PRS_index, function(x) dim(x)) # resulting in this many individuals

# turn list into dataframe
# pivot to long format with column 'type' containing plusminus, proportion, PAFGRS, PAFGRSplus, FLS, PRS
indicators_PRS_index <- do.call(rbind.data.frame, indicators_PRS_index) |>
    pivot_longer(cols = -c(pident, coding, DEP_index), names_to = "indicator", values_to = "value") |>
    mutate(indicator = factor(indicator, levels = c("plusminus", "proportion", "FLS", "PAFGRS", "PAFGRSplus", "PRS")))

# Exclude related index individuals
# Identify individuals based on family relatedness and case-control status
pident_to_exclude <- indicators_PRS_index |>
    pivot_wider(
        names_from = c(coding, indicator),
        values_from = c(value)
    ) |>
    mutate(fam_number = NESDA_related$famnumber.x[match(pident, NESDA_related$pident)]) |>
    filter(!is.na(fam_number)) |>
    group_by(fam_number) |>
    mutate(n = n()) |>
    filter(n > 1) |>
    select(pident, DEP_index, fam_number, n) |>
    arrange(fam_number) |>
    mutate(exclude = dplyr::coalesce(DEP_index == 1 & any(DEP_index == 0, na.rm = TRUE), FALSE)) |>
    mutate(
        # If more than one row is not excluded, randomly exclude all but one
        extra_exclude = if (sum(!exclude, na.rm = TRUE) > 1) {
            temp <- which(!exclude)
            to_exclude <- sample(temp, length(temp) - 1)
            seq_along(exclude) %in% to_exclude
        } else {
            FALSE
        }
    ) |>
    mutate(final_exclude = exclude | extra_exclude)
pident_to_exclude

# Sanity check: comparison with those excluded in acylcarnitines project
pident_to_exclude$pident %in% c(
    110025, 110015, 110104, 210042, 330029, 330003, 330055, 330370, 330060, 330303,
    130003, 330112, 110282, 330143, 330142, 330377, 330103, 330009, 310012, 330298,
    210127, 330113, 330197, 110799, 210178, 330369, 210172, 330246, 330116, 220122,
    210242, 330358, 210272, 330228, 210283, 330295, 110701, 330334, 330388, 330330,
    330356, 330355, 330354, 210458, 320033, 210305, 210303, 330302, 210529, 210466,
    220310, 219493, 210509, 310598
)

# Exclude these individuals from indicators_PRS_index
indicators_PRS_index <- indicators_PRS_index |>
    filter(!(pident %in% pident_to_exclude$pident[pident_to_exclude$final_exclude]))
```


# Save

```{r}
# Save indicators
save(indicators_with_index_dep, file = here("Int_results", "indicators_with_index_dep.RData"))
save(indicators_with_index_anx, file = here("Int_results", "indicators_with_index_anx.RData"))

pident_for_FH <- FH$pident
save(pident_for_FH, file = here("Int_results", "pident_for_FH.RData"))

# Save pidents that should be exluded due to relatedness
save(pident_to_exclude, file = here("Int_results", "pident_to_exclude.RData"))

# Save
save(FH, file = here("Int_results", "FH.RData"))
#save(indicators_PRS_index, file = here("Int_results", "indicators_PRS_index.RData"))
#save(indicators_with_index, file = here("Int_results", "indicators_with_index.RData"))
```