---
title: "LT-FH Simu DEP/ANX/BIP/SCZ"
output: html_document
date: "2024-04-05"
---

This script
1) Simulates families with different structures and disorders and assesses their prediction by family history indicators
2) Compares proportion vs. dense proportion family history indicators
3) Assesses prediction accuracy (AUC) for depression, schizophrenia and bipolar disorder
4) Checks if PA-FGRS gives the same results as LT-FH
Input: Scripts/setup.R, Scripts/indicator_functions.R
Output: Results/AUC_DEP_denseprop.pdf, Results/AUC_SCZ.pdf, Results/AUC_BIP.pdf, Results/bip_prediction_output.RData, Results/scz_prediction_output.RData, Results/dep_prediction_denseprop.RData
Status: DEBUG (problem with full_sim function)

```{r setup, include=FALSE}
# Setup - packages and functions
source(file = "setup.R")

# # Build the family
# num_sibs <- 2 # number of siblings to simulate
# parents <- c("f", "m") # mother and father
# fam_members <- c("index", paste0("sib", sequence(num_sibs)), parents)

# General notes
# Ks <- c(DEP =  0.19, ANX = 0.2, SCZ = 0.005, BIP = 0.013) # prevalence in Netherlands from NEMESIS-2
# h2ls <- c(DEP =  0.4, ANX = 0.4, SCZ = 0.77, BIP = 0.67) # twin heritability estimates from Polderman et al.
rg_SCZ.BIP.DEP.ANX <- matrix(c( # ldsc regression genetic correlation estimates from Grotzinger et al.
1, 0.682, 0.357, 0.367,
0.682, 1, 0.343, 0.297,
0.357, 0.343, 1, 0.867,
0.367, 0.357, 0.867, 1
),4,4)
rownames(rg_SCZ.BIP.DEP.ANX) <- colnames(rg_SCZ.BIP.DEP.ANX) <- c("SCZ", "BIP", "DEP", "ANX")


rg_DEP.ANX <- rg_SCZ.BIP.DEP.ANX[c("DEP", "ANX"), c("DEP", "ANX")]
rg_SCZ.BIP <- rg_SCZ.BIP.DEP.ANX[c("SCZ", "BIP"), c("SCZ", "BIP")]
rg_DEP.ANX.BIP <- rg_SCZ.BIP.DEP.ANX[c("DEP", "ANX", "BIP"), c("DEP", "ANX", "BIP")]

sim_reps <- 10
sim_size <- 1e4 # number of simulated persons in one simulation round (for rmvn)
```


## Compare prop & dense prop (predicting depression)

```{r DEP, warning=FALSE}
sim.input.list <- list(
# Depression 
  list(name="DEP, 1 fam member", index_disorder = "DEP", Ks = c(DEP = 0.19), h2ls = c(DEP = 0.4), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1"))
  ,list(name="DEP, 3 fam members", index_disorder = "DEP", Ks = c(DEP = 0.19), h2ls = c(DEP = 0.4), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "m", "f"))
  ,list(name="DEP/ANX, 3 fam members", index_disorder = "DEP", Ks = c(DEP = 0.19, ANX = 0.2), h2ls = c(DEP = 0.4, ANX = 0.4), rg = rg_DEP.ANX, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "m", "f"))
  ,list(name="DEP/ANX/BIP, 3 fam members", index_disorder = "DEP", Ks = c(DEP =  0.19, ANX = 0.2, BIP = 0.013), h2ls = c(DEP =  0.4, ANX = 0.4, BIP = 0.67), rg = rg_DEP.ANX.BIP, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "m", "f"))
  ,list(name="DEP/ANX/BIP/SCZ, 3 fam members", index_disorder = "DEP", Ks = c(DEP =  0.19, ANX = 0.2, SCZ = 0.005, BIP = 0.013), h2ls = c(DEP =  0.4, ANX = 0.4, SCZ = 0.77, BIP = 0.67), rg = rg_SCZ.BIP.DEP.ANX, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "m", "f"))
  ,list(name="DEP/ANX/BIP/SCZ, 8 fam members", index_disorder = "DEP", Ks = c(DEP =  0.19, ANX = 0.2, SCZ = 0.005, BIP = 0.013), h2ls = c(DEP =  0.4, ANX = 0.4, SCZ = 0.77, BIP = 0.67), rg = rg_SCZ.BIP.DEP.ANX, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "sib3", "sib4", "sib5", "sib6", "m", "f"))
)

dep_prediction <- full_sim(sim.input.list = sim.input.list)
#save(dep_prediction, file = here("Results", "dep_prediction_denseprop.RData"))

means_se <- as.data.frame(do.call("rbind", map(dep_prediction, "outcome_meanse")))
means_se

plusminus_AUC <- as.data.frame(do.call("rbind",
lapply(map(dep_prediction, "outcome_meanse"), # all listed dfs
       function(x) x[,"FH_DEP_lm.status.plusminus_AUC"] ) # extract same column
))
plusminus_AUC

proportion_AUC <- as.data.frame(do.call("rbind",
lapply(map(dep_prediction, "outcome_meanse"), # all listed dfs
       function(x) x[,"FH_DEP_lm.status.proportion_AUC"] ) # extract same column
))
proportion_AUC

denseprop_AUC <- as.data.frame(do.call("rbind",
lapply(map(dep_prediction, "outcome_meanse"), # all listed dfs
       function(x) x[,"FH_DEP_lm.status.denseprop_AUC"] ) # extract same column
))
denseprop_AUC

ltfh_AUC <- as.data.frame(do.call("rbind",
lapply(map(dep_prediction, "outcome_meanse"), # all listed dfs
       function(x) x[,"LTFH_DEP_lm.status.muliab_AUC"] ) # extract same column
))
ltfh_AUC
```


```{r denseprop.plot}
x <- unlist(map(sim.input.list, "name"))
# Plot R2 by family size for depression
AUC_plot_denseprop <- ggplot() +
  # plusminus
  geom_point(data = plusminus_AUC, aes(x = x, y = mean, color = "FH+-"), position = position_nudge(-.33), size = 3) +
  geom_errorbar(data = plusminus_AUC, aes(x = x, ymin = mean - se, ymax = mean + se, color = "FH+-"), width = .2, position = position_nudge(-.33)) +
  # denseprop
  geom_point(data = denseprop_AUC, aes(x = x, y = mean, color = "Prop Dense"), position = position_nudge(-.11), size = 3) +
  geom_errorbar(data = denseprop_AUC, aes(x = x, ymin = mean - se, ymax = mean + se, color = "Prop Dense"), width = .2, position = position_nudge(-.11)) +
  # proportion
  geom_point(data = proportion_AUC, aes(x = x, y = mean, color = "FH Proportion"), position = position_nudge(.11), size = 3) +
  geom_errorbar(data = proportion_AUC, aes(x = x, ymin = mean - se, ymax = mean + se, color = "FH Proportion"), width = .2, position = position_nudge(.11)) +
  # ltfh
  geom_point(data = ltfh_AUC, aes(x = x, y = mean, color = "LT-FH"), position = position_nudge(.33), size = 3) +
  geom_errorbar(data = ltfh_AUC, aes(x = x, ymin = mean - se, ymax = mean + se, color = "LT-FH"), width = .2, position = position_nudge(.33)) +
  
  scale_color_manual(values = c("FH+-" = "#4B0C6BFF", "Prop Dense" = "#00BFFF", "FH Proportion" = "#CF4446FF", "LT-FH"= "#FB9A06FF"),
                     labels = c("FH Proportion", "FH+-", "LT-FH", "Prop Dense"),
                     guide = guide_legend(
                       title = "Predictor" 
                     )) +
  ylim(0.5, 0.7) +
  ylab("AUC") +
  xlab("Family") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, size = 12, vjust = 1, hjust = 1),
        axis.text.y = element_text(size = 12),
        title = element_text(size = 15),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12)) +
  ggtitle("Assessment of Depression Prediction", subtitle = "AUC of true.status & predictor")

AUC_plot_denseprop
ggsave(here("Results", "AUC_DEP_denseprop.pdf"), AUC_plot_denseprop, width = 20, height = 20, units = "cm")
```

This is to compare whether proportion of possible diagnoses (here FH Proportion) or proportion of family members affected with _any_ disorder (here Prop Dense) is the better predictor. These two indicators are equivalent when only one disorder is taken into account, but not when multiple diagnoses are considered. Therefore, we specifically plot various diagnostic combinations with disorders relevant for the prediction of depression.


## Predicting Schizophrenia
```{r warning = FALSE}
sim.input.list <- list(
# Schizophrenia
  list(name="SCZ, 1 fam member", index_disorder = "SCZ", Ks = c(SCZ = 0.005), h2ls = c(SCZ = 0.77), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1"))
  
  ,list(name="SCZ, 3 fam members", index_disorder = "SCZ", Ks = c(SCZ = 0.005), h2ls = c(SCZ = 0.77), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "m", "f"))
  
  ,list(name="SCZ/BIP, 3 fam members", index_disorder = "SCZ", Ks = c(SCZ = 0.005, BIP = 0.013), h2ls = c(SCZ = 0.77, BIP = 0.67), rg = rg_SCZ.BIP, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "m", "f"))
  
  ,list(name="SCZ/DEP/ANX/BIP, 3 fam members", index_disorder = "SCZ", Ks = c(DEP =  0.19, ANX = 0.2, SCZ = 0.005, BIP = 0.013), h2ls = c(DEP =  0.4, ANX = 0.4, SCZ = 0.77, BIP = 0.67), rg = rg_SCZ.BIP.DEP.ANX, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "m", "f"))
  
  ,list(name="SCZ, 8 fam members", index_disorder = "SCZ", Ks = c(SCZ = 0.005), h2ls = c(SCZ = 0.77), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "sib3", "sib4", "sib5", "sib6", "m", "f"))
  
  ,list(name="SCZ/BIP, 8 fam members", index_disorder = "SCZ", Ks = c(SCZ = 0.005, BIP = 0.013), h2ls = c(SCZ = 0.77, BIP = 0.67), rg = rg_SCZ.BIP, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "sib3", "sib4", "sib5", "sib6", "m", "f"))
  
  ,list(name="SCZ/DEP/ANX/BIP, 8 fam members", index_disorder = "SCZ", Ks = c(DEP =  0.19, ANX = 0.2, SCZ = 0.005, BIP = 0.013), h2ls = c(DEP =  0.4, ANX = 0.4, SCZ = 0.77, BIP = 0.67), rg = rg_SCZ.BIP.DEP.ANX, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "sib3", "sib4", "sib5", "sib6", "m", "f"))
)

scz_prediction <- full_sim(sim.input.list = sim.input.list)

save(scz_prediction, file = here("Results", "scz_prediction_output.RData"))
#load(here("Results", "scz_prediction_output.RData"))

means_se <- as.data.frame(do.call("rbind", map(scz_prediction, "outcome_meanse")))
means_se

plusminus_AUC <- as.data.frame(do.call("rbind",
lapply(map(scz_prediction, "outcome_meanse"), # all listed dfs
       function(x) x[,"FH_SCZ_lm.status.plusminus_AUC"] ) # extract same column
))
plusminus_AUC

proportion_AUC <- as.data.frame(do.call("rbind",
lapply(map(scz_prediction, "outcome_meanse"), # all listed dfs
       function(x) x[,"FH_SCZ_lm.status.proportion_AUC"] ) # extract same column
))
proportion_AUC

# denseprop_AUC <- as.data.frame(do.call("rbind",
# lapply(map(scz_prediction, "outcome_meanse"), # all listed dfs
#        function(x) x[,"FH_SCZ_lm.status.denseprop_AUC"] ) # extract same column
# ))
# denseprop_AUC

ltfh_AUC <- as.data.frame(do.call("rbind",
lapply(map(scz_prediction, "outcome_meanse"), # all listed dfs
       function(x) x[,"LTFH_SCZ_lm.status.muliab_AUC"] ) # extract same column
))
ltfh_AUC
```


```{r}
# Plot R2 by disorder number for depression
x <- unlist(map(sim.input.list, "name"))
# Plot R2 by family size for depression
AUC_plot_SCZ <- ggplot() +
  # plusminus
  geom_point(data = plusminus_AUC, aes(x = x, y = mean, color = "FH+-"), position = position_nudge(-.22), size = 3) +
  geom_errorbar(data = plusminus_AUC, aes(x = x, ymin = mean - se, ymax = mean + se, color = "FH+-"), width = .2, position = position_nudge(-.22)) +
  # proportion
  geom_point(data = proportion_AUC, aes(x = x, y = mean, color = "FH Proportion"), position = position_nudge(0), size = 3) +
  geom_errorbar(data = proportion_AUC, aes(x = x, ymin = mean - se, ymax = mean + se, color = "FH Proportion"), width = .2, position = position_nudge(0)) +
  # ltfh
  geom_point(data = ltfh_AUC, aes(x = x, y = mean, color = "LT-FH"), position = position_nudge(.22), size = 3) +
  geom_errorbar(data = ltfh_AUC, aes(x = x, ymin = mean - se, ymax = mean + se, color = "LT-FH"), width = .2, position = position_nudge(.22)) +
  
  scale_color_manual(values = c("FH+-" = "#4B0C6BFF", "FH Proportion" = "#CF4446FF", "LT-FH"= "#FB9A06FF"),
                     labels = c( "FH Proportion", "FH +-", "LT-FH"),
                     guide = guide_legend(
                       title = "Predictor"
                     )) +
  ylim(0.5, 0.8) +
  ylab("AUC") +
  xlab("Family") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, size = 12, vjust = 1, hjust = 1),
        axis.text.y = element_text(size = 12),
        title = element_text(size = 15),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12)) +
  ggtitle("Assessment of Schizophrenia Prediction", subtitle = "AUC of true.status & predictor")

AUC_plot_SCZ
ggsave(here("Results", "AUC_SCZ.pdf"), AUC_plot_SCZ, width = 20, height = 20, units = "cm")
```




## Predicting Bipolar
```{r warning = FALSE}
sim.input.list <- list(
# Bipolar disorder
  list(name="BIP, 1 fam member", index_disorder = "BIP", Ks = c(BIP = 0.013), h2ls = c(BIP = 0.67), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1"))
  
  ,list(name="BIP, 3 fam members", index_disorder = "BIP", Ks = c(BIP = 0.013), h2ls = c(BIP = 0.67), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "m", "f"))
  
  ,list(name="BIP/SCZ, 3 fam members", index_disorder = "BIP", Ks = c(SCZ = 0.005, BIP = 0.013), h2ls = c(SCZ = 0.77, BIP = 0.67), rg = rg_SCZ.BIP, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "m", "f"))
  
  ,list(name="BIP/DEP/ANX/SCZ, 3 fam members", index_disorder = "BIP", Ks = c(DEP =  0.19, ANX = 0.2, SCZ = 0.005, BIP = 0.013), h2ls = c(DEP =  0.4, ANX = 0.4, SCZ = 0.77, BIP = 0.67), rg = rg_SCZ.BIP.DEP.ANX, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "m", "f"))
  
  ,list(name="BIP, 8 fam members", index_disorder = "BIP", Ks = c(BIP = 0.013), h2ls = c(BIP = 0.67), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "sib3", "sib4", "sib5", "sib6", "m", "f"))
  
  ,list(name="BIP/SCZ, 8 fam members", index_disorder = "BIP", Ks = c(SCZ = 0.005, BIP = 0.013), h2ls = c(SCZ = 0.77, BIP = 0.67), rg = rg_SCZ.BIP, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "sib3", "sib4", "sib5", "sib6", "m", "f"))
  
  ,list(name="BIP/DEP/ANX/SCZ, 8 fam members", index_disorder = "BIP", Ks = c(DEP =  0.19, ANX = 0.2, SCZ = 0.005, BIP = 0.013), h2ls = c(DEP =  0.4, ANX = 0.4, SCZ = 0.77, BIP = 0.67), rg = rg_SCZ.BIP.DEP.ANX, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "sib3", "sib4", "sib5", "sib6", "m", "f"))
)

bip_prediction <- full_sim(sim.input.list = sim.input.list)

save(bip_prediction, file = here("Results", "bip_prediction_output.RData"))
#load(here("Results", "bip_prediction_output.RData"))

means_se <- as.data.frame(do.call("rbind", map(bip_prediction, "outcome_meanse")))
means_se

plusminus_AUC <- as.data.frame(do.call("rbind",
lapply(map(bip_prediction, "outcome_meanse"), # all listed dfs
       function(x) x[,"FH_BIP_lm.status.plusminus_AUC"] ) # extract same column
))
plusminus_AUC

proportion_AUC <- as.data.frame(do.call("rbind",
lapply(map(bip_prediction, "outcome_meanse"), # all listed dfs
       function(x) x[,"FH_BIP_lm.status.proportion_AUC"] ) # extract same column
))
proportion_AUC

# denseprop_AUC <- as.data.frame(do.call("rbind",
# lapply(map(bip_prediction, "outcome_meanse"), # all listed dfs
#        function(x) x[,"FH_BIP_lm.status.denseprop_AUC"] ) # extract same column
# ))
# denseprop_AUC

ltfh_AUC <- as.data.frame(do.call("rbind",
lapply(map(bip_prediction, "outcome_meanse"), # all listed dfs
       function(x) x[,"LTFH_BIP_lm.status.muliab_AUC"] ) # extract same column
))
ltfh_AUC
```


```{r}
# Plot AUC for bipolar
x <- unlist(map(sim.input.list, "name"))
# Plot R2 by family size for depression
AUC_plot_BIP <- ggplot() +
  # plusminus
  geom_point(data = plusminus_AUC, aes(x = x, y = mean, color = "FH+-"), position = position_nudge(-.22), size = 3) +
  geom_errorbar(data = plusminus_AUC, aes(x = x, ymin = mean - se, ymax = mean + se, color = "FH+-"), width = .2, position = position_nudge(-.22)) +
  # proportion
  geom_point(data = proportion_AUC, aes(x = x, y = mean, color = "FH Proportion"), position = position_nudge(0), size = 3) +
  geom_errorbar(data = proportion_AUC, aes(x = x, ymin = mean - se, ymax = mean + se, color = "FH Proportion"), width = .2, position = position_nudge(0)) +
  # ltfh
  geom_point(data = ltfh_AUC, aes(x = x, y = mean, color = "LT-FH"), position = position_nudge(.22), size = 3) +
  geom_errorbar(data = ltfh_AUC, aes(x = x, ymin = mean - se, ymax = mean + se, color = "LT-FH"), width = .2, position = position_nudge(.22)) +
  
  scale_color_manual(values = c("FH+-" = "#4B0C6BFF", "FH Proportion" = "#CF4446FF", "LT-FH"= "#FB9A06FF"),
                     labels = c( "FH Proportion", "FH +-", "LT-FH"),
                     guide = guide_legend(
                       title = "Predictor"
                     )) +
  ylim(0.5, 0.8) +
  ylab("AUC") +
  xlab("Family") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, size = 12, vjust = 1, hjust = 1),
        axis.text.y = element_text(size = 12),
        title = element_text(size = 15),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12)) +
  ggtitle("Assessment of Bipolar Disorder Prediction", subtitle = "AUC of true.status & predictor")

AUC_plot_BIP
ggsave(here("Results", "AUC_BIP.pdf"), AUC_plot_BIP, width = 20, height = 20, units = "cm")
```


# Checking PA-FGRS

```{r}
sim.input.list <- list(
# Depression 
  list(name="DEP, 4 fam members", index_disorder = "DEP", Ks = c(DEP = 0.19), h2ls = c(DEP = 0.4), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "m", "f"))
  ,list(name="DEP, 8 fam members", index_disorder = "DEP", Ks = c(DEP = 0.19), h2ls = c(DEP = 0.4), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "sib3", "sib4", "sib5", "sib6", "m", "f"))
  # Schizphrenia
  ,list(name="SCZ, 4 fam members", index_disorder = "SCZ", Ks = c(SCZ = 0.005), h2ls = c(SCZ = 0.77), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "m", "f"))
  ,list(name="SCZ, 8 fam members", index_disorder = "SCZ", Ks = c(SCZ = 0.005), h2ls = c(SCZ = 0.77), rg = 1, sim_reps = sim_reps, sim_size = sim_size, fam_members = c("index", "sib1", "sib2", "sib3", "sib4", "sib5", "sib6", "m", "f"))  
)

test <- full_sim(sim.input.list = sim.input.list)
dep_test <- list(test[[1]], test[[2]])

means_se <- as.data.frame(do.call("rbind", map(dep_test, "outcome_meanse")))
means_se


proportion_AUC <- as.data.frame(do.call("rbind",
lapply(map(dep_test, "outcome_meanse"), # all listed dfs
       function(x) x[,"FH_DEP_lm.status.proportion_AUC"] ) # extract same column
))
proportion_AUC

ltfh_AUC <- as.data.frame(do.call("rbind",
lapply(map(dep_test, "outcome_meanse"), # all listed dfs
       function(x) x[,"LTFH_DEP_lm.status.muliab_AUC"] ) # extract same column
))
ltfh_AUC

pafgrs_AUC <- as.data.frame(do.call("rbind",
lapply(map(dep_test, "outcome_meanse"), # all listed dfs
       function(x) x[,"PAFGRS_DEP_lm.status.muliab_AUC"] ) # extract same column
))
pafgrs_AUC
```

```{r}
scz_test <- list(test[[3]], test[[4]])

proportion_AUC <- as.data.frame(do.call("rbind",
lapply(map(scz_test, "outcome_meanse"), # all listed dfs
       function(x) x[,"FH_SCZ_lm.status.proportion_AUC"] ) # extract same column
))
proportion_AUC

ltfh_AUC <- as.data.frame(do.call("rbind",
lapply(map(scz_test, "outcome_meanse"), # all listed dfs
       function(x) x[,"LTFH_SCZ_lm.status.muliab_AUC"] ) # extract same column
))
ltfh_AUC

pafgrs_AUC <- as.data.frame(do.call("rbind",
lapply(map(scz_test, "outcome_meanse"), # all listed dfs
       function(x) x[,"PAFGRS_SCZ_lm.status.muliab_AUC"] ) # extract same column
))
pafgrs_AUC
```

