---
title: "Prediction of Depression by FH - Visualisation"
output: html_document
date: "2024-12-09"
---

This script creates the folllowing visualizations:
1) Descriptives of FTI data availability, also by assessment type
2) Prevalences of disorders in relatives
3) Descriptives of final sample (Table 1)
4) Plot of sex- and age-specific prevalences for PAFGRS+Sex/Age
5) Generation of full correlation matrices and heatmaps to display (Figure 3)
6) Distributions of indicators (Figure 2)
7) AUC plots (Figure 4)
8) OR plots (Figure 5)
Input: Scripts/indicator_functions.R, assessment.RData, indicators_PRS_index.RData, AUC_results_combined.RData, assessment_display4.RData, NESDA_DAP2406/N1_A100R.sav, FTI_famsize.csv, FH.RData
Output: Main and supplementary figures and tables (see final code chunk)
Status: Running

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Clear workspace
rm(list = ls())
getwd()

rmarkdown::pandoc_available()
# Setup - packages and functions
source(file = "Scripts/indicator_functions.R")

# Load the FH indicators (output of calculate_indicators.Rmd and assess_prediction.Rmd)
load(file = here("Results", "assessment.RData")) # assessment
load(file = here("Int_results", "indicators_PRS_index.RData")) # indicators_PRS_index. each df contains pident, index, PRS, respective indicators
load(file = here("Int_results", "AUC_results_combined.RData")) # AUC and AUC.SE for combined prediction
load(file = here("Int_results", "assessment_display4.RData")) # dataframe for display item
load(file = here("Int_results", "results_df_scaled.RData")) # results with ORs per SD increase
# NESDA baseline descriptives
NESDA_descriptives <- read_sav(file = here("Data", "NESDA_DAP2406", "N1_A100R.sav"))
FTI_famsize <- read.csv(here("Data", "FTI_famsize.csv"))
load(file = here("Int_results", "FH.RData")) # family history

# Turn DEP_index into factor
indicators_PRS_index <- indicators_PRS_index |>
  mutate(DEP_index = as.factor(DEP_index))
sum(is.na(indicators_PRS_index$value))
head(indicators_PRS_index)
dim(indicators_PRS_index)
sum(!duplicated(indicators_PRS_index$pident))

# Dataframe for certain descriptives
# Reshape, extract relevant columns, keep only pidents that are in indicators_PRS_index (equivalent to final sample)
FH_long <- FH |>
  pivot_longer(
    cols = -c(pident, DEP_index), # adjust to match all relevant columns
    names_to = "key",
    values_to = "value"
  ) |>
  extract(
    col = key,
    into = c("disorder", "coding", "imputation", "fammember"),
    regex = "^([^_]+)_([^_]+)(?:_([^_]+))?_([^_]+)$", # specifies chunks in column names, with 3rd one being optional
    remove = TRUE
  ) |>
  filter(pident %in% indicators_PRS_index$pident) # keep only pidents that are present in indicators_PRS_index

head(FH_long)
dim(FH_long)
sum(!duplicated(FH_long$pident))
assessment
```

# FTI descriptives: Number of relatives and such
```{r}
# Number of individuals for each coding type
simples <- FH_long |>
  filter(coding == "simple", imputation == "", disorder == "depdiag") |>
  group_by(pident) |>
  summarise(
    n = sum(!is.na(value))
  ) |>
  count(n)

elaborates <- FH_long |>
  filter(coding == "elaborate", imputation == "", disorder == "depdiag") |>
  group_by(pident) |>
  summarise(
    n = sum(!is.na(value))
  ) |>
  count(n)

simples_imputed <- FH_long |>
  filter(coding == "simple", imputation == "0imputed", disorder == "depdiag") |>
  group_by(pident) |>
  summarise(
    n = sum(!is.na(value))
  ) |>
  count(n)

elaborates_imputed <- FH_long |>
  filter(coding == "elaborate", imputation == "0imputed", disorder == "depdiag") |>
  group_by(pident) |>
  summarise(
    n = sum(!is.na(value))
  ) |>
  count(n)

# How many individuals have how many relatives with FH data?
p1 <- simples %>%
  ggplot(aes(x = factor(n), y = nn)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Number of relatives", y = "Count") +
  ggtitle("Short assessment") +
  theme_minimal()

p2 <- elaborates %>%
  ggplot(aes(x = factor(n), y = nn)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Number of relatives", y = "Count") +
  ggtitle("Extended assessment") +
  theme_minimal()

p3 <- simples_imputed %>%
  ggplot(aes(x = factor(n), y = nn)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Number of relatives", y = "Count") +
  ggtitle("Short assessment, imputed") +
  theme_minimal()

p4 <- elaborates_imputed %>%
  ggplot(aes(x = factor(n), y = nn)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Number of relatives", y = "Count") +
  ggtitle("Extended assessment, imputed") +
  theme_minimal()

p1 + p2 + p3 + p4 +
  plot_annotation(
    title = "Number of relatives with FH data",
    theme = theme(plot.title = element_text(hjust = 0.5))
  )

# Supplementary Table X
# How many individuals are left with no relatives with FH data? Only in the non-imputed
simple_n <- c(sum(simples$nn[simples$n != 0]), sum(simples_imputed$nn[simples_imputed$n != 0]))
elaborates_n <- c(sum(elaborates$nn[elaborates$n != 0]), sum(elaborates_imputed$nn[elaborates_imputed$n != 0]))
gain_in_index <- as.data.frame(rbind(simple_n, elaborates_n))
colnames(gain_in_index) <- c("No imputation", "With imputation")
rownames(gain_in_index) <- c("Short assessment", "Extended assessment")
gain_in_index
``` 

# Exploring prevalences and sample sizes

```{r} 
info <- FH_long |>
  filter(coding != "2020paper") |>
  group_by(imputation, coding, disorder) |>
  summarise(
    n = sum(!is.na(value)),
    n_cases = sum(value, na.rm = TRUE),
    prevalence = sum(value, na.rm = TRUE) / sum(!is.na(value)), .groups = "drop"
  ) |>
  bind_rows(
    FH_long |>
      filter(coding != "2020paper") |>
      group_by(DEP_index, imputation, coding, disorder) |>
      summarise(
        n = sum(!is.na(value)),
        n_cases = sum(value, na.rm = TRUE),
        prevalence = sum(value, na.rm = TRUE) / sum(!is.na(value)), .groups = "drop"
      )
  )
info

# Supplementary Table X
# Effect of imputation (gain in index individuals see above)
# Depression
dep_imp <- info |>
  filter(is.na(DEP_index), disorder == "depdiag") |>
  pivot_wider(
    names_from = c(imputation),
    values_from = c(prevalence, n)
  ) |>
  select(-DEP_index)|> # switch rows to reverse
  arrange(desc(coding))
# Anxiety
anx_imp <- info |>
  filter(is.na(DEP_index), disorder == "anxdiag") |>
  pivot_wider(
    names_from = c(imputation),
    values_from = c(prevalence, n)
  ) |>
  select(-DEP_index) |> # switch rows to reverse
  arrange(desc(coding))
imputation_effect <- as.data.frame(rbind(dep_imp, anx_imp))

# Supplementary Table X
# Crosstable simple/elaborate
temp2 <- FH_long |>
  pivot_wider(
    names_from = c(coding, imputation, disorder),
    values_from = c(value)
  )
head(temp2)
table(temp2$simple__depdiag, temp2$elaborate__depdiag, useNA = "always")
table(temp2$fammember, temp2$simple__depdiag, useNA = "always")


# Supplementary Table X - Manually filled, print values here
# Contrasting short and extended assessment
# get the number of individuals that don't have count == 0
FH_long |>
  filter(coding == "simple", imputation == "", disorder == "depdiag") |>
  group_by(DEP_index, pident) |>
  summarise(
    n = sum(!is.na(value))
  ) |>
  count(n != 0)

FH_long |>
  filter(coding == "elaborate", imputation == "", disorder == "depdiag") |>
  group_by(DEP_index, pident) |>
  summarise(
    n = sum(!is.na(value))
  ) |>
  count(n != 0)

# Prevalence in cases versus control
info |>
  filter(!is.na(DEP_index), disorder == "depdiag", imputation == "") |>
  select(-imputation, -n) |>
  pivot_wider(
    names_from = c(DEP_index),
    values_from = c(prevalence)
  )

info |>
  filter(!is.na(DEP_index), disorder == "anxdiag", imputation == "") |>
  select(-imputation, -n) |>
  pivot_wider(
    names_from = c(DEP_index),
    values_from = c(prevalence)
  )

# FH+ percentages in cases versus controls
indicators_PRS_index |>
  filter(coding == "DEP_simple", indicator == "plusminus") |>
  group_by(DEP_index) |>
  summarise(
    n_plus = sum(value, na.rm = TRUE),
    n_total = sum(!is.na(value)),
    percentage = n_plus / n_total
  )

indicators_PRS_index |>
  filter(coding == "DEP_elaborate", indicator == "plusminus") |>
  group_by(DEP_index) |>
  summarise(
    n_plus = sum(value, na.rm = TRUE),
    n_total = sum(!is.na(value)),
    percentage = n_plus / n_total
  )

assessment


# Correlation PAFGRS and PRS
indicators_temp <- indicators_PRS_index |>
  filter(coding == "DEP_simple") |>
  pivot_wider(names_from = indicator, values_from = value)
cor.test(indicators_df$PAFGRS, indicators_df$PRS, method = "pearson", use = "pairwise.complete.obs")
indicators_temp <- indicators_PRS_index |>
  filter(coding == "DEP_elaborate") |>
  pivot_wider(names_from = indicator, values_from = value)
cor.test(indicators_df$PAFGRS, indicators_df$PRS, method = "pearson", use = "pairwise.complete.obs")
```

# Descriptives of final sample - Table 1
```{r}
# Make dataframe with all relevant variables
data_for_descriptives <- indicators_PRS_index |>
  filter(coding == "DEP_simple") |>
  pivot_wider(
    names_from = indicator,
    values_from = value
  ) |>
  left_join(NESDA_descriptives, by = "pident") |>
  left_join(FTI_famsize, by = "pident")
head(data_for_descriptives)
# and add one indicator that includes ANX
data_for_descriptives <- indicators_PRS_index |>
  filter(coding == "DEP_simple_with_ANX_simple") |>
  pivot_wider(
    names_from = indicator,
    values_from = value
  ) |>
  mutate(PAFGRSplusANX = PAFGRSplus) |>
  select(pident, PAFGRSplusANX) |>
  left_join(data_for_descriptives, by = "pident")

# Summarise
sample <- data_for_descriptives |>
  group_by(DEP_index) |>
  summarise(
    n = n(),
    mean_age = mean(Age, na.rm = TRUE),
    sd_age = sd(Age, na.rm = TRUE),
    percentage_female = sum(Sexe == 2, na.rm = TRUE) / n(),
    sib_nr_median = median(sib_nr, na.rm = TRUE),
    sib_nr_iqr = paste(quantile(sib_nr, 0.25, na.rm = TRUE) + 2,
      quantile(sib_nr, 0.75, na.rm = TRUE) + 2,
      sep = "-"
    ),
    sib_nr_range = paste(range(sib_nr, na.rm = TRUE) + 2, collapse = "-"),
    proportion_plusminus = sum(plusminus, na.rm = TRUE) / n(),
    mean_proportion = mean(proportion, na.rm = TRUE),
    sd_proportion = sd(proportion, na.rm = TRUE),
    mean_FLS = mean(FLS, na.rm = TRUE),
    sd_FLS = sd(FLS, na.rm = TRUE),
    mean_PAFGRS = mean(PAFGRS, na.rm = TRUE),
    sd_PAFGRS = sd(PAFGRS, na.rm = TRUE),
    mean_PAFGRSplus = mean(PAFGRSplus, na.rm = TRUE),
    sd_PAFGRSplus = sd(PAFGRSplus, na.rm = TRUE),
    mean_PAFGRSplusANX = mean(PAFGRSplusANX, na.rm = TRUE),
    sd_PAFGRSplusANX = sd(PAFGRSplusANX, na.rm = TRUE),
    mean_PRS = mean(PRS, na.rm = TRUE),
    sd_PRS = sd(PRS, na.rm = TRUE)
  )
View(sample)

# Build Table 1
table_1 <- sample |>
  mutate(
    age = sprintf("%.2f (%.2f)", mean_age, sd_age),
    percentage_female = sprintf("%.2f", percentage_female * 100),
    sib_nr_summary = sprintf("%s (%s; %s)", sib_nr_median + 2, sib_nr_iqr, sib_nr_range),
    plusminus = sprintf("%.2f", proportion_plusminus * 100),
    proportion = sprintf("%.2f (%.2f)", mean_proportion, sd_proportion),
    FLS = sprintf("%.2f (%.2f)", mean_FLS, sd_FLS),
    PAFGRS = sprintf("%.2f (%.2f)", mean_PAFGRS, sd_PAFGRS),
    PAFGRSplus = sprintf("%.2f (%.2f)", mean_PAFGRSplus, sd_PAFGRSplus),
    PAFGRSplusANX = sprintf("%.2f (%.2f)", mean_PAFGRSplusANX, sd_PAFGRSplusANX),
    PRS = sprintf("%.2f (%.2f)", mean_PRS, sd_PRS)
  ) |>
  select(
    n, age, percentage_female, sib_nr_summary,
    plusminus, proportion, FLS, PAFGRS, PAFGRSplus, PAFGRSplusANX, PRS
  )
table_1 <- as.data.frame(t(table_1))
rownames(table_1) <- c(
  "N", "Age (Mean [SD])", "Sex (% female)", "Number of 1st degree relatives (Median [IQR; Range])",
  "FH+/- (% FH +)", "Proportion (Mean [SD])", "FLS (Mean [SD])",
  "PAFGRS (Mean [SD])", "PAFGRS+Sex/Age (Mean [SD])", "PAFGRS+Sex/Age/ANX (Mean [SD])", "PRS (Mean [SD])"
)
colnames(table_1) <- c("Controls", "MDD Cases")
View(table_1)

# b) 1st degree relatives with DEP or ANX or both
# how many relatives have both depression and anxiety?
temp <- FH_long |>
  filter(coding == "simple" & imputation == "") |>
  dplyr::select(pident, disorder, fammember, value) |>
  pivot_wider(
    id_cols = c(pident, fammember),
    names_from = disorder,
    values_from = value
  )

# number of relatives with both dep and anx
both <- temp |>
  filter(depdiag == 1 & anxdiag == 1) |>
  dim()
# number of relatives with info on either dep or anx or both
complete <- temp |>
  filter(!is.na(depdiag) | !is.na(anxdiag)) |>
  dim()
# Percentage of those with both disorders from the total relatives
both / complete
```


# Sex- and age-specific prevalences
```{r}
K_info
# reshape to long format
K_long <- K_info %>%
  pivot_longer(cols = -ages, names_to = "group", values_to = "value")

# nicer labels for legend
K_long <- K_long %>%
  mutate(group_label = recode(group,
    DEP_female = "Major Depressive Disorder: Female",
    DEP_male   = "Major Depressive Disorder: Male",
    ANX_female = "Any Anxiety Disorder: Female",
    ANX_male   = "Any Anxiety Disorder: Male"
  ))

# define colors
cols <- c(
  "Major Depressive Disorder: Female" = "#3ecae2",
  "Major Depressive Disorder: Male"   = "#2bb1e6",
  "Any Anxiety Disorder: Female" = "#d48032",
  "Any Anxiety Disorder: Male"   = "#c2470e"
)

# create plot
prevalence_model <- ggplot(K_long, aes(x = ages, y = value, color = group_label, linetype = group_label)) +
  geom_line(size = 1.1) +
  geom_point(size = 2) +
  scale_color_manual(values = cols) +
  scale_linetype_manual(values = c("solid", "solid", "dashed", "dashed")) +
  scale_x_continuous(breaks = K_info$ages) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) + # show as percentages
  labs(
    x = "Age",
    y = "Cumulative incidence",
    color = "Group",
    linetype = "Group"
  ) +
  theme_minimal(base_size = 9) +
  theme(
    plot.title = element_text(size = 11, face = "bold", hjust = 0.5),
    legend.position = "right"
  )

# print plot
print(prevalence_model)

```


# Correlations
```{r}
cors_pearson <- list()
cors_spearman <- list()

for (type in all_types) {
  indicators_df <- indicators_PRS_index |>
    filter(coding == type) |>
    pivot_wider(names_from = indicator, values_from = value)

  cors_pearson[[type]] <- cov2cor(cov(indicators_df[, predictors], method = "pearson", use = "na.or.complete"))
  cors_spearman[[type]] <- cov2cor(cov(indicators_df[, predictors], method = "spearman", use = "na.or.complete"))
}

cors_pearson
cors_spearman
save(cors_pearson, cors_spearman, file = here("Int_results", "cors.RData"))

# Draw selected heatmaps
draw_heatmap(round(cors_pearson[["DEP_simple"]], 3), title = "Pearson Correlation", subtitle = "Depression prediction from FH")

cors_pearson
cors_pearson[["DEP_elaborate"]]
```

## Cor matrix df prep
```{r}
indicators_ANX_simple <- indicators_PRS_index |>
  filter(coding == "DEP_simple_with_ANX_simple") |>
  pivot_wider(names_from = indicator, values_from = value) |>
  rename("plusminus+ANX" = "plusminus", "proportion+ANX" = "proportion", "FLS+ANX" = "FLS", "PAFGRS+ANX" = "PAFGRS", "PAFGRSplus+ANX" = "PAFGRSplus")

indicators_DEP_simple <- indicators_PRS_index |>
  filter(coding == "DEP_simple") |>
  pivot_wider(names_from = indicator, values_from = value)

all_indicators <- cbind(indicators_DEP_simple[, 4:8], indicators_ANX_simple[, 4:9]) |>
  rename("FH+/-" = "plusminus", "Proportion" = "proportion", "PAFGRS+Sex/Age" = "PAFGRSplus", "PAFGRS+Sex/Age/Anx" = "PAFGRSplus+ANX", "PGS" = "PRS")

glimpse(all_indicators)
```

## Display item 3 - Pearson's r
```{r}
cors <- cov2cor(cov(all_indicators, use = "na.or.complete", method = "pearson"))

# Alternative using cor.test()
cor.test(all_indicators$Proportion, all_indicators$PAFGRS, use = "pairwise", method = "pearson", adjust = "none")
glimpse(all_indicators)
p_values <- matrix(NA, ncol = ncol(all_indicators), nrow = ncol(all_indicators))
colnames(p_values) <- colnames(all_indicators)
rownames(p_values) <- colnames(all_indicators)
for (i in 1:ncol(all_indicators)) {
  for (j in 1:ncol(all_indicators)) {
    if (i != j) {
      p_values[i, j] <- cor.test(all_indicators[, i], all_indicators[, j], use = "pairwise", method = "pearson")$conf[2]
    }
  }
}
p_values

p_values <- matrix(NA, ncol = ncol(all_indicators), nrow = ncol(all_indicators))
colnames(p_values) <- colnames(all_indicators)
rownames(p_values) <- colnames(all_indicators)
for (i in 1:ncol(all_indicators)) {
  for (j in 1:ncol(all_indicators)) {
    if (i != j) {
      p_values[i, j] <- cor_test_against_value(all_indicators[, i], all_indicators[, j], rho0 = 0.9)$p_value
    }
  }
}
View(p_values)
# Note: SE is always 0.026 because it depends only on the sample size

cors
# Display item 3
cors_df <- cors[
  c("FH+/-", "Proportion", "FLS", "PAFGRS", "PAFGRS+Sex/Age", "PAFGRS+Sex/Age/Anx", "PGS"),
  c("FH+/-", "Proportion", "FLS", "PAFGRS", "PAFGRS+Sex/Age", "PAFGRS+Sex/Age/Anx", "PGS")
]

# Draw selected heatmaps
heatmap_pearson <- draw_heatmap(cors_df,
  subtitle = "Family History Indicators and PGS",
  title = "Correlations between genetic predictors for depression",
  legend_title = "Pearson's r"
)
heatmap_pearson
```


## Cor heatmap - Spearman's rank rho
```{r}
cors <- cov2cor(cov(all_indicators, use = "na.or.complete", method = "spearman"))

cors_df <- cors[
  c("FH+/-", "Proportion", "FLS", "PAFGRS", "PAFGRS+Sex/Age", "PAFGRS+Sex/Age/Anx", "PRS"),
  c("FH+/-", "Proportion", "FLS", "PAFGRS", "PAFGRS+Sex/Age", "PAFGRS+Sex/Age/Anx", "PRS")
]
cors_df

# Draw selected heatmaps
heatmap_spearman <- draw_heatmap(cors_df,
  subtitle = "Family History Indicators and PRS",
  title = "Correlations between genetic predictors for depression",
  legend_title = "Spearman's rho"
)
heatmap_spearman
```


# Indicator distributions

```{r}
# Initiate empty list to store plots
viz <- list()

fill_colors <- c(
  "0" = scales::alpha("#3482A4FF", 1),
  "1" = scales::alpha("#A4E0BBFF", 0.6)
)

# Define theme
my_theme <- theme_minimal() +
  theme(
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 9),
    axis.text.x = element_text(size = 9),
    axis.text.y = element_text(size = 9),
    legend.position = "none"
  )

for (type in all_types) {
  indicators_df <- indicators_PRS_index |>
    filter(coding == type) |>
    pivot_wider(names_from = indicator, values_from = value)

  # Distribution of plusminus
  plusminus <- indicators_df |>
    summarise(
      N_plus = sum(plusminus, na.rm = TRUE),
      N_minus = (sum(!is.na(plusminus)) - sum(plusminus, na.rm = TRUE))
    ) |>
    bind_rows(
      indicators_df |>
        group_by(DEP_index) |>
        summarise(
          N_plus = sum(plusminus, na.rm = TRUE),
          N_minus = (sum(!is.na(plusminus)) - sum(plusminus, na.rm = TRUE))
        )
    )

  # check overlap in distribution of indicators between cases and controls
  density_prop <- indicators_df |>
    ggplot(aes(x = proportion, fill = DEP_index)) +
    geom_density(position = "identity", adjust = 0.5) +
    labs(
      x = "Proportion",
      y = "Density",
      fill = "Status of Index"
    ) +
    scale_fill_manual(
      values = fill_colors, # Colors for groups
      labels = c("0" = "Controls", "1" = "Dep.Cases") # Custom labels
    ) +
    my_theme

  density_prop

  density_FLS <- indicators_df %>%
    ggplot(aes(x = FLS, fill = DEP_index)) +
    geom_density(position = "identity", adjust = 0.5) +
    labs(
      x = "FLS",
      y = "Density",
      fill = "Status of Index"
    ) +
    scale_fill_manual(
      values = fill_colors, # Colors for groups
      labels = c("0" = "Controls", "1" = "Dep.Cases") # Custom labels
    ) +
    my_theme

  density_PAFGRS <- indicators_df %>%
    ggplot(aes(x = PAFGRS, fill = DEP_index)) +
    geom_density(position = "identity", adjust = 0.5) +
    labs(
      x = "PAFGRS",
      y = "Density",
      fill = "Status of Index"
    ) +
    scale_y_continuous(limits = c(0, 3)) +
    scale_fill_manual(
      values = fill_colors, # Colors for groups
      labels = c("0" = "Controls", "1" = "Dep.Cases") # Custom labels
    ) +
    my_theme

  density_PAFGRSplus <- indicators_df %>%
    ggplot(aes(x = PAFGRSplus, fill = DEP_index)) +
    geom_density(position = "identity", adjust = 0.5) +
    labs(
      x = "PAFGRS+Sex/Age",
      y = "Density",
      fill = "Status of Index"
    ) +
    scale_y_continuous(limits = c(0, 3)) +
    scale_fill_manual(
      values = fill_colors, # Colors for groups
      labels = c("0" = "Controls", "1" = "Dep.Cases") # Custom labels
    ) +
    my_theme

  density_PRS <- indicators_df %>%
    ggplot(aes(x = PRS, fill = DEP_index)) +
    geom_density(position = "identity", adjust = 1.5) +
    labs(
      x = "PGS",
      y = "Density",
      fill = "Status of Index"
    ) +
    scale_y_continuous(limits = c(0, 0.5)) +
    scale_fill_manual(
      values = fill_colors, # Colors for groups
      labels = c("0" = "Controls", "1" = "Dep.Cases") # Custom labels
    ) +
    my_theme


  # Plot assembly
  density_assembly <- ((density_prop + density_FLS) / (density_PAFGRS + density_PAFGRSplus)) +
    plot_layout(guides = "collect") +
    plot_annotation(
      title = paste0("FH Indicators for Depression", " (", type, ")"),
      theme = theme(
        plot.title = element_text(size = 40, face = "bold")
      )
    ) & theme_minimal()

  viz[[type]] <- list(
    plusminus = plusminus,
    density_prop = density_prop, density_FLS = density_FLS,
    density_PAFGRS = density_PAFGRS, density_PAFGRSplus = density_PAFGRSplus,
    density_PRS = density_PRS,
    density_assembly = density_assembly
  )
}


# extract all density_assembly plots
density <- map(viz, "density_assembly")
#density
# save(viz, file = here("Results", "NESDA_DEP_viz.RData"))


# make barplot in ggplot of plus over plus + minus
plusminus <- viz$DEP_simple$plusminus %>%
  filter(!is.na(DEP_index)) %>%
  ggplot(aes(x = DEP_index, y = N_plus / (N_plus + N_minus), fill = DEP_index)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  ylim(0, 1) +
  labs(
    y = "Proportion of FH+",
    fill = "Status of Index"
  ) +
  scale_fill_manual(
    values = fill_colors, # Colors for groups
    labels = c("0" = "Controls", "1" = "MDD Cases") # Custom labels
  ) +
  my_theme +
  guides(fill = "none")

plusminus <- plusminus + theme_minimal() + theme(axis.title.x = element_blank())
proportion <- viz$DEP_simple$density_prop + theme_minimal()
FLS <- viz$DEP_simple$density_FLS + theme_minimal() + theme(axis.title.y = element_blank())
PAFGRS <- viz$DEP_simple$density_PAFGRS + labs(x = "PAFGRS") + theme_minimal()
PAFGRS_2 <- viz$DEP_simple$density_PAFGRSplus + theme_minimal() + theme(axis.title.y = element_blank())
PAFGRS_3 <- viz$DEP_simple_with_ANX_simple$density_PAFGRSplus + labs(x = "PAFGRS+Sex/Age/ANX") + theme_minimal() + theme(axis.title.y = element_blank())
PRS <- viz$DEP_simple$density_PRS + theme_minimal() + theme(axis.title.y = element_blank())

proportion

# Display item 2
layout <- "
ABCD
EFGH
"

display_2 <- plusminus + proportion + FLS + guide_area() +
  PAFGRS + PAFGRS_2 + PAFGRS_3 + PRS +
  plot_layout(design = layout, guides = "collect") 
display_2
```


# AUC plots

```{r}
# List of AUC plots
predictors_supplement <- c("FH+/-", "Proportion", "FLS", "PAFGRS", "PAFGRS+Sex/Age", "PGS")
predictors <- factor(predictors_supplement, levels = predictors_supplement)
AUC_plots <- list()
for (type in types) {
  AUC_df <- assessment[[type]] |>
    t() |>
    as.data.frame() |>
    rename(AUC = AUC, SE = AUC.SE)
  AUC_plots[[type]] <- plot_AUC(AUC_df, groups = predictors)
}
AUC_plots

# With Anxiety as helper trait
predictors_ANX <- c("plusminus+ANX", "proportion+ANX", "FLS+ANX", "PAFGRS+ANX", "PAFGRSplus+ANX", "PGS")
# List of AUC plots
predictors_ANX <- factor(predictors_ANX, levels = predictors_ANX)
AUC_plots_ANX <- list()
for (type in types_ANX) {
  AUC_df <- assessment[[type]] |>
    t() |>
    as.data.frame() |>
    rename(AUC = AUC, SE = AUC.SE)
  AUC_plots_ANX[[type]] <- plot_AUC(AUC_df, groups = predictors_ANX, title = type)
}
AUC_plots_ANX

# Supplemental Figure: AUC plots assembly
plot1 <- AUC_plots$DEP_simple + ggtitle("Short assessment")
plot2 <- AUC_plots$DEP_elaborate + ggtitle("Extended assessment")
plot3 <- AUC_plots$DEP_simple_0imputed + ggtitle("Short assessment + Imputation")
plot4 <- AUC_plots$DEP_elaborate_0imputed + ggtitle("Extended assessment + Imputation")
assessment_assembly <- (plot1 + plot2) / (plot3 + plot4)
assessment_assembly
```

```{r }
# Display item 4
assessment_display4

predictors_4a <- c("FH+/-", "Proportion", "FLS", "PAFGRS", "PAFGRS+Sex/Age", "PAFGRS+Sex/Age/ANX", "PGS", "PAFGRS+Sex/Age/ANX\n& PGS")
predictors_4a <- factor(predictors_4a, levels = predictors_4a)
predictors_4a

display_4 <- ggplot(data = assessment_display4[1:8, ]) +
  aes(x = predictors_4a, y = AUC) +
  geom_errorbar(aes(x = predictors_4a, ymin = AUC - SE, ymax = AUC + SE), colour = "#35A0ABFF", width = .2, position = position_nudge(0), size = 1) +
  geom_point(colour = "#35A0ABFF", size = 9 / .pt, shape = 18, position = position_nudge(0)) +
  ylim(0.50, 0.75) +
  ylab("AUC") +
  xlab("") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, size = 9, vjust = 1, hjust = 1),
    axis.text.y = element_text(size = 9)
  )
display_4
```


```{r }
# Display item 4b
results_df_scaled

predictors_4b <- c("FH+/-", "Proportion", "FLS", "PAFGRS", "PAFGRS+Sex/Age", "PAFGRS+Sex/Age/ANX", "PGS", "PAFGRS+Sex/Age/ANX\nwith PGS", "PGS \nwith PAFGRS+Sex/Age/ANX")
predictors_4b <- factor(predictors_4b, levels = predictors_4b)
predictors_4b

display_4b <- ggplot(data = results_df_scaled[1:9, ]) +
  aes(x = predictors_4b, y = OR) +
  geom_errorbar(aes(x = predictors_4b, ymin = ci_lower, ymax = ci_upper), colour = "#35A0ABFF", width = .2, position = position_nudge(0), size = 1) +
  geom_point(colour = "#35A0ABFF", size = 9 / .pt, shape = 18, position = position_nudge(0)) +
  ylim(1, 3) +
  ylab("Odds Ratio (per SD increase)") +
  xlab("") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, size = 9, vjust = 1, hjust = 1),
    axis.text.y = element_text(size = 9)
  )
display_4b
```

# OR plot

```{r}
# Based on script by Pietro
PGS_PAFGRS_df <- indicators_PRS_index |>
  filter(coding == "DEP_simple_with_ANX_simple") |>
  pivot_wider(names_from = indicator, values_from = value) |>
  dplyr::select(DEP_index, PAFGRSplus, PRS) |>
  filter(!is.na(PRS))

df <- PGS_PAFGRS_df
variable <- "PRS" # Change to "PAFGRS" for PAFGRS analysis
num_quantiles <- 5 # Number of quantiles to create

## Plot: Display item 4b
pgs_results_quantiles <- logistic_regression_quantiles(PGS_PAFGRS_df, "PRS", num_quantiles = 5) # You can change the num_quantiles to your desired value
pafgrs_results_quantiles <- logistic_regression_quantiles(PGS_PAFGRS_df, "PAFGRSplus", num_quantiles = 5)
# Table for supplement
head(pgs_results_quantiles)
quantile_details <- cbind(pgs_results_quantiles$quantile, pafgrs_results_quantiles$se, pafgrs_results_quantiles$case_nr, pafgrs_results_quantiles$control_nr,
  pgs_results_quantiles$se, pgs_results_quantiles$case_nr, pgs_results_quantiles$control_nr)

# Combine the results into one data frame
combined_results_quantiles <- bind_rows(pgs_results_quantiles, pafgrs_results_quantiles)
combined_results_quantiles

display_4c <- ggplot(combined_results_quantiles) +
  aes(x = quantile, y = odds_ratio, color = variable) +
  geom_point(size = 9 / .pt, shape = 18, position = position_dodge(0.22)) + # Dodge points horizontally
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2, position = position_dodge(0.22), size = 1) + # Thicker error bars
  labs(
    x = "Quantile", y = "Odds Ratio", color = "Genetic Indicator"
  ) +
  scale_color_manual(
    values = c("PRS" = "#3A2C59FF", "PAFGRSplus" = "#5ACCADFF"),
    labels = c("PRS" = "PGS", "PAFGRSplus" = "PAFGRS+Sex/Age/ANX")
  ) +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 9),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9)
  )
display_4c
```

# Save figures and tables
```{r}
# Main
write.csv(table_1, file = here("Results", "Table_1.csv"), row.names = TRUE)
ggsave(here("Results", "Figure3.pdf"), heatmap_pearson, width = 21, height = 12.7, units = "cm")
ggsave(here("Results", "Figure2.pdf"), display_2, width = 21, height = 12.7, units = "cm")
ggsave(here("Results", "Figure4a.pdf"), display_4, width = 21, height = 12.7, units = "cm")
ggsave(here("Results", "Figure4b.pdf"), display_4b, width = 21, height = 12.7, units = "cm")
ggsave(here("Results", "Figure4c.pdf"), display_4c, width = 21, height = 12.7, units = "cm")

# Supplement
ggsave(here("Results", "cor_heatmap_spearman.pdf"), heatmap_spearman, width = 21, height = 12.7, units = "cm")
ggsave(here("Results", "Supplemental_Figure_AUC_assessment.pdf"), assessment_assembly, width = 21, height = 16, units = "cm")
write.csv(quantile_details, file = here("Results", "quantile_details.csv"), row.names = FALSE)
ggsave(here("Results", "Age-sex-specific_prevalences.pdf"), prevalence_model, width = 21, height = 12.7, units = "cm")
write.csv(gain_in_index, file = here("Results", "gain_in_index.csv"), row.names = TRUE)
write.csv(imputation_effect, file = here("Results", "imputation_effect.csv"), row.names = TRUE)
```